Supporting Information : Complex and informative dependencies propogate throughout ecology and can be revealed using copulas

### 2019-05-13 SupportingInformation.Rmd
more edits

# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and similar 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of Q1
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of dependence. Thus this comparison helps answer the
third part of Q1 for bivariate 
datasets.

@@

<!--table for Best-fitting copulas, green spruce aphid abundance data-->
```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, green spruce aphid abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

@@

<!--table for Best-fitting copulas, leaf plum curling aphid first flight data-->
```{r tab_ff_bestcop, echo=F, results="markup"}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data. Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

@@

<!--table for Best-fitting copulas, Ceratium furca abundance data-->
```{r tab_pns_bestcop, echo=F, results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
             format='latex', linesep = "",
             caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
             booktabs=TRUE
             )
```

@@

<!--table for Best-fitting copulas, methane-flux data-->
```{r tab_methane_bestcop, echo=F, results="markup"}
knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, methane-flux data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
             booktabs=TRUE
             )
```

@@

# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
N-copula null hypothesis, and answering question Q1 from the Introduction.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

<!--Plots for AR(1) with Clayton-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model,
C copula.]{Comparison of copula structure of noise inputs and model 
outputs, AR(1) model, C copula. Red and blue points give 
means over 50 replicate simulations of the statistics listed on the left axes, 
red points are values for noise inputs and blue points are for
model outputs. Error bars are standard errors. 
$P$-values (purple points, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions of 50 noise and 
output values have the same mean; $p$-values less than $0.05$ indicate that 
copula structure as measured by the listed statistic differs significant
between noise inputs and model outputs. See methods section for Q2 
for additional details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}

@@

<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
selection procedures detailed in the 
methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

@@

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have N-copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```

@@

# From causes to consequences for green spruce aphid abundance data\label{sect:greenspruce}
Describe the winter temp data
Show the copulas with winter temp, 10 locations

@@

>>>>

# Verifying model selection results of testing the observational hypothesis on bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and similar 
Kendall (or Spearman)
correlation. Thus the comparison addresses the first two parts of the observational hypothesis
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of dependence. Thus this comparison helps address the
third part of the observational hypothesis for bivariate 
datasets.

@@

<!--table for Best-fitting copulas, green spruce aphid abundance data-->
```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, green spruce aphid abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for the observational hypothesis. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

@@

<!--table for Best-fitting copulas, leaf plum curling aphid first flight data-->
```{r tab_ff_bestcop, echo=F, results="markup"}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data. Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for the observational hypothesis. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

@@

<!--table for Best-fitting copulas, Ceratium furca abundance data-->
```{r tab_pns_bestcop, echo=F, results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
             format='latex', linesep = "",
             caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data. See methods for the observational hypothesis for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
             booktabs=TRUE
             )
```

@@

<!--table for Best-fitting copulas, methane-flux data-->
```{r tab_methane_bestcop, echo=F, results="markup"}
knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, methane-flux data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data. See methods for the observational hypothesis for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
             booktabs=TRUE
             )
```

@@

# Verifying model selection results of testing the observational hypothesis on multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
N-copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

<!--Plots for AR(1) with Clayton-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model,
C copula.]{Comparison of copula structure of noise inputs and model 
outputs, AR(1) model, C copula. Red and blue points give 
means over 50 replicate simulations of the statistics listed on the left axes, 
red points are values for noise inputs and blue points are for
model outputs. Error bars are standard errors. 
$P$-values (purple points, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions of 50 noise and 
output values have the same mean; $p$-values less than $0.05$ indicate that 
copula structure as measured by the listed statistic differs significant
between noise inputs and model outputs. See methods section for the causal hypotheses 
for additional details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}

@@

<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the second model detailed in 
the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
selection procedures detailed in the 
methods section for the observational hypothesis. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

@@

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the second model detailed in 
the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for the observational hypothesis. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have N-copula structure
(see methods for the observational hypothesis). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")
```

@@

# From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
Winter temperature for year $t$ was for December of year $t-1$ to March of 
year $t$. Data, which pertained to the same locations and years as the aphid data,
were obtained and processed exactly as in [@Sheppard2016]. Copulas
with winter temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
copulas that paralleled the analyses applied to bivariate ecological datasets 
(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.

@@

<!--***DAN: Shya, please insert here a table like Table 2 in your recent doc "explore_aphid_abund_wintertemp_dep". Please label it tab:greenspruce. Note that 
currently there are two NA values in the first column, but there should be none.-->
